---
title: "EDA and Collaborative Filtering"
author: "Brandon Hoeft"
date: "October 22, 2017"
output: 
  html_document: 
    keep_md: yes # http://happygitwithr.com/rmd-test-drive.html#output-format
    toc: yes
    toc_depth: 3
---

```{r setup, include = FALSE}
# my global options defined for each code chunk.
knitr::opts_chunk$set(fig.width=8, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, comment = '')
```

## Introduction

This Markdown file has most of my code, relevant plots, and explanations of decisions made so far in the data wrangling process.

I will update it with collaborative filtering model information too, as needed. 

## Getting the data into R

I used the following packages to do most of the initial data pre-processing and exploratory analysis. I'm working on an AWS AMI instance that's a linux machine with RStudio and git installed, using one of the instances publicly available and pre-configured from [Louis Aslett's website](http://www.louisaslett.com/RStudio_AMI/). I initially started with the free tier, *t2.micro* instance, but once I started the modeling phase, I scaled up to *m4.large*, which is also a general purpose computing instance but with better RAM, CPU, and better optimized memory management. AWS AMI instance costs can be calculated [here](https://calculator.s3.amazonaws.com/index.html). 

The datasets queried by Matt Hayden, I've put in my own S3 Bucket on Amazon Web Services (AWS). As such, the `aws.s3` [package](https://github.com/cloudyr/aws.s3) was used to access that bucket and bring the datasets into my environment. Access Keys are needed to make read/write calls to the bucket, but I've intentionally left that code out of here. 

``` {r echo = TRUE}

library(aws.s3) 
library(readr) # faster alternatives to base read. methods. 
library(lubridate) # date wrangling. 
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # data viz. 
library(pryr) # mem_used() and object_size() functions to manage/understand memory usage.
```

To read the Part items data and the user webactivity data from s3 I passed read_table2() functions for reading tabular data to read items dataset and the read_csv() function to read in the web user click data CSV files. 

``` {r eval = FALSE, include = FALSE}
# specify personal account keys as environment variables so I can read my s3 object(s) from AWS. 

# DO NOT DO NOT DO NOT DO NOT DO NOT SAVE KEY in code or render in output!!!! Could compromise AWS account. 
Sys.setenv("AWS_ACCESS_KEY_ID" = "",
           "AWS_SECRET_ACCESS_KEY" = "")

```

``` {r echo = TRUE}

# items: part number, parent, catalogue, attributes/values.
items <- s3read_using(FUN = read_table2, 
                      object = "obfuscatedItems_10_17_17.txt", 
                      col_names = TRUE,
                      col_types = "cciciiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiii",
                      bucket = "pred498team5")

# users: click data from the company website for a random day of user's selected, their activity for past 3 months and click summaries of how they interacted with parts. 

# web user click data from Feb - Mar 2017
users_febmar17 <- s3read_using(FUN = read_csv, 
                      col_names = TRUE,
                      col_types = "ccici",
                      object = "obfuscatedWebActivity7124.csv", 
                      bucket = "pred498team5")

# web user click data from April - June 2017
users_aprmayjun17 <- s3read_using(FUN = read_csv, 
                      col_names = TRUE,
                      col_types = "ccici",
                      object = "obfuscatedWebActivity7127.csv", 
                      bucket = "pred498team5")

# web user click data from July - August 2017
users_julaug17 <- s3read_using(FUN = read_csv, 
                      col_names = TRUE,
                      col_types = "ccici",
                      object = "obfuscatedWebActivity7129.csv", 
                      bucket = "pred498team5")


```

The 3 separate user click activity queries performed by Matt on his company's graph database were required in order to get around maximum data request requirements. They contain the same information, with the following columns:

``` {r echo = TRUE}
names(users_aprmayjun17)
```

These 3 datasets of user click activity covering different time periods between February - August 2017 need to be unioned, or combined, into a single dataset. 

``` {r echo = TRUE}

users <- bind_rows(users_febmar17, users_aprmayjun17, users_julaug17)

```

## Wrangling the Data

The **items** dataset contains [obfuscated](https://en.wikipedia.org/wiki/Obfuscation_(software)) information about all of the PartNumbers in the company's inventory, as well as information on the catalog that the PartNumber may have been listed in, and the first 20 attributes and attribute values describing the specifications of each PartNumber. It is a very large dataset with `r nrow(items)` rows, representing the unique PartNumbers.  

The only information we're particularly interested in from the **items** dataset is the Parent column, and left joining that to the matching PartNumber in the users data. Building a recommender system using the Parent family instead of the specific PartNumber will be discussed later. The items dataset has a lot of columns that describe attributes of each PartNumber and Parent. For purposes of a collaborative filtering model, we don't need to know these, any columns that include *"Attr"*, *"Val"*, or *"Catalog"* in its name are dropped. 

``` {r echo = TRUE}
user_items <- users %>%
  left_join(items, by = "PartNumber") %>%
  mutate(ActionDate = ymd(ActionDate), # parse into a date format.
         ActionId_label = factor(ActionId, # create labels
                                 labels = c("add to order", "select Part",
                                            "select Part detail",
                                            "print detail",
                                            "save CAD drawing detail",
                                            "print CAD drawing detail"))) %>% 
  select(-starts_with("Val"), -starts_with("Attr"), -starts_with("Catalog"))
```

Check that the left join preserved all of the original data in our full users dataset after joining in the Parent value associated with each PartNumber searched by users.

``` {r echo = TRUE}
all.equal(users[1:5], user_items[1:5])
```

Our original datasets are taking up a lot of memory. Specifically, all other datasets other than user_items are currently taking up `r object_size(users, items, users_febmar17, users_aprmayjun17, users_julaug17)`MB of RAM. We'll remove them since everything we need in terms of user click data, the PartNumbers, Parent, actions taken, and the action date are all in the user_items dataset. Our main user_items dataset is quite large, requiring `r object_size(user_items)` MB of RAM. 

``` {r echo = TRUE}
# We can remove the other datasets we don't need anymore to save available RAM. 
remove(users, items, users_febmar17, users_aprmayjun17, users_julaug17)

```


### High Level Description of the User & Items Data

The user click data (**may refer to users interchangeably as visitors**) came from three separate graph database queries that were required in order to get around maximum data request requirements. Together, the 3 user datasets represent click data from a random sample of visitors to the company's website who have an account with the company. These visitors indvidual session click activities were queried over a time period from February 2017 through August 2017. These data describe how users/visitors of a manufacturing supplier are interacting with the company's website. Sessions within the context of this dataset refers to a specific date that the user engaged with a PartNumber. 

With a baseline dataset, I can start describing some information about the user activity. 

We'll start by asking some questions of the data. The web activity covered by these data begins on `r min(user_items$ActionDate)` and goes through `r max(user_items$ActionDate) `, covering a span of `r max(user_items$ActionDate) - min(user_items$ActionDate)` days. 

``` {r echo = TRUE}
user_items %>%
  group_by(ActionDate) %>%
  ggplot()

```

#### Summary of Unique Visitors, PartNumbers, Parents

Initial analysis shows that there are `r unique(user_items$VisitorId) %>% length()` unique VisitorID numbers in this dataset. 

There are `r unique(user_items$PartNumber) %>% length()` unique PartNumbers associated with the visitor web activity data.

Instead of looking at very specific PartNumbers, we can also roll the analysis up to the Parent level. The Parent level is a grouping of similar PartNumber products to a parent family. Within the website, it's a hyperlink associated with the PartNumber. Parent groups will vary. It is a way that individual merchandising managers at the company decide to categorize and split stuff up. There are `r unique(user_items$Parent) %>% length()` unique Parent family numbers (referred herein as Parents) associated with the visitor web activity data. 

Code for getting the aforementioned details is below. 

``` {r echo = TRUE, eval = FALSE }
# How many Different Users are in the dataset? 
unique(user_items$VisitorId) %>% length()
# How many distinct PartNumber's are in the dataset? 
unique(user_items$PartNumber) %>% length()
# How many distinct Parent Family's of PartNumbers are in the dataset? 
unique(user_items$Parent) %>% length()

```

#### How do total Action Counts of each visitor distribute? 

Action counts are how often a visitor did a particular action within their session. In a single session (the day of the visitor activity interaction with a PartNumber) if they selected a PartNumber in the carousel, this is reflected as a single row in the dataset, with an ActionCount of 1. If they selected the part 2 times in this session, the ActionCount value is equal to 2. 

``` {r echo = TRUE}
user_items %>%
  group_by(VisitorId) %>%
  summarize(activity_count = sum(ActionCount)) %>%
  with(summary(activity_count))

user_items %>%
  group_by(VisitorId) %>%
  summarize(activity_count = sum(ActionCount)) %>%
  ggplot(aes(activity_count, y = ..density..)) + 
  geom_histogram(binwidth = 100, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(activity_count)),
             color = "dodgerblue3", linetype = "dashed", size = 0.5) +
  geom_text(aes(0,.002, family = "courier", label = paste("median = ", median(activity_count))),
            nudge_x = 700, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Total interactions with PartNumbers by each Visitor',
       subtitle = 'binwidth of 100 ',
       x = "Total Actions per Visitor")

```

After careful consideration, it was determined that total actions per PartNumber or per Parent should not influence the implicit rating derived for the recommender. This could be for a variety of reasons, such as selecting a part multiple times in the same session may perhaps be just as indicative of uncertainty as it could be of endorsing that part. 

#### How many unique PartNumbers is each visitor interacting with? 


``` {r echo = TRUE}

user_items %>%
  group_by(VisitorId) %>%
  distinct(PartNumber) %>% # only keep distinct part numbers per visitor.
  summarize(unique_part_count = n()) %>% # get count within the group. 
  with(summary(unique_part_count))

user_items %>%
  group_by(VisitorId) %>%
  distinct(PartNumber) %>% # only keep distinct part numbers per visitor.
  summarize(unique_part_count = n()) %>% # get count within the group. 
  ggplot(aes(unique_part_count, y = ..density..)) + 
  geom_histogram(binwidth = 50, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(unique_part_count)),
             color = "dodgerblue3", linetype = "dashed", size = 0.5) +
  geom_text(aes(0,.004, family = "courier", label = paste("median = ", median(unique_part_count))), 
            nudge_x = 250, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Unique PartNumber Interactions by each Visitor',
       subtitle = 'binwidth of 50',
       x = 'Unique PartNumbers per Visitor')

```


#### How many unique Parents are visitors interacting with? 

Instead of looking at the distinct PartNumbers that each visitor interacts with, we can also look at the Parent category that each PartNumber rolls up to. By reducing the diversity of different products from the specific PartNumber to the Parent category they roll up to, we'd expect the count of interactions with unique Parents to decrease. However, this may improve the sparsity problem of a recommender system. 

``` {r echo = TRUE}

user_items %>%
  group_by(VisitorId) %>%
  distinct(Parent) %>% # only keep distinct parent per visitor.
  summarize(unique_parent_count = n()) %>% # get count within the group. 
  with(summary(unique_parent_count))

user_items %>%
  group_by(VisitorId) %>%
  distinct(Parent) %>% # only keep distinct parent per visitor.
  summarize(unique_parent_count = n()) %>% # get count within the group. 
  ggplot(aes(unique_parent_count, y = ..density..)) + 
  geom_histogram(binwidth = 50, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(unique_parent_count)),
             color = "dodgerblue3", linetype = "dashed", size = 0.5) +
  geom_text(aes(0,.004, family = "courier", label = paste("median = ", median(unique_parent_count))), 
            nudge_x = 150, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Unique Parent Interactions',
       subtitle = 'binwidth of 50',
       x = 'Unique Parent category per Visitor')
```


#### How many unique visitors have interacted with each PartNumber? 

``` {r echo = TRUE}
user_items %>%
  group_by(PartNumber) %>%
  distinct(VisitorId) %>% # only keep distinct visitors per Item.
  summarize(unique_users_count = n()) %>%
  with(summary(unique_users_count))

user_items %>%
  group_by(PartNumber) %>%
  distinct(VisitorId) %>% # only keep distinct visitors per Item.
  summarize(unique_users_count = n()) %>%
  ggplot(aes(unique_users_count, y = ..density..)) + # modal value is around log(2) or 2 items.
  geom_histogram(binwidth = 5, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(unique_users_count)),
             color = "dodgerblue3", linetype = "dashed", size = 0.5) +
  geom_text(aes(0,.1, family = "courier", label = paste("median = ", median(unique_users_count))), 
            nudge_x = 50, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Unique Visitor Interactions for each PartNumber',
       subtitle = 'binwidth of 5',
       x = 'Unique Visitors per PartNumber')
```

#### How many unique visitors have interacted with each Parent Family? 

Because Parent Family's are a broader generalization than PartNumbers, we'd 
expect to have more interactions within Parent Family than at the very specific PartNumber level and the summary statistics and distributions reflect that. 

``` {r echo = TRUE}
user_items %>%
  group_by(Parent) %>%
  distinct(VisitorId) %>% # only keep distinct visitors per Item.
  summarize(unique_users_count = n()) %>%
  with(summary(unique_users_count))

user_items %>%
  group_by(Parent) %>%
  distinct(VisitorId) %>% # only keep distinct visitors per Item.
  summarize(unique_users_count = n()) %>%
  ggplot(aes(unique_users_count, y = ..density..)) + 
  geom_histogram(binwidth = 20, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(unique_users_count)),
             color = "dodgerblue2", linetype = "dashed", size = 0.5) +
  geom_text(aes(0, .02, family = "courier", label = paste("median = ", median(unique_users_count))),
            nudge_x = 150, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Unique Visitors for each Parent Family',
       subtitle = 'binwidth of 20',
       x = 'Unique Visitors per PartNumber')

```

#### How many PartNumbers fall into each Parent Family?

It's also interesting to know when we reduce the dimensionality of the products by rolling up from PartNumber to their parent categories, how many PartNumbers make up the Parent? 

``` {r echo = TRUE}
user_items %>%
  group_by(Parent) %>%
  distinct(PartNumber) %>% # remove duplicate part numbers within Parent Group. 
  summarize(unique_partnumber_count = n()) %>% # return count of unique items per parent. 
  with(summary(unique_partnumber_count))
```

#### (TO EXPLORE) Do Parents with more PartNumbers also have more overall visitor clicks? 

IF we roll PartNumbers up to the parent level, there may be a potential concern that the frequency of ratings per a Parent category may be a function of the diversity of PartNumbers that make up the Parent. If this is the case, we'd likely see a correlation pattern between the frequency of PartNumbers per Parent and the number of unique visitors per Parent. In other words, the most popular parents could just be the ones with the most PartNumbers in these data.

That could be a potential source of concern or area for further research in these data as collaborative filtering recommendations can be influenced by the most popular items.

#### What types of click Actions do the visitors take? 

Below is the distribution of the 6 different click actions in the visitor click dataset of `r nrow(user_items)` different click actions. 

``` {r echo = TRUE}
user_items %>%
  group_by(ActionId, ActionId_label) %>%
  summarize(frequency = n()) %>%
  arrange(desc(frequency))

# Bar graph of the 6 different Action ID Labels
user_items %>%
  group_by(ActionId, ActionId_label) %>%
  summarize(frequency = n()) %>%
  ggplot(aes(x = reorder(ActionId_label, desc(frequency)), y = frequency / sum(frequency))) +
    geom_bar(aes(fill = ActionId_label), stat = "identity") +
    labs(title = "Distribution of click Actions taken by Users",
         subtitle = "on company website or mobile app",
         x = NULL,
         y = "Percent") +
    geom_text(aes(label = frequency, family = "courier", ), 
              size = 3,
              stat= "identity", 
              vjust = -.5) +
    guides(fill=FALSE) + # remove legend
    theme(axis.text.x=element_text(angle=25,hjust=1)) + # tilt x labels
    scale_y_continuous(labels=scales::percent)

```

#### (TO EXPLORE) What percent of the time is selecting a part a precursor to adding to cart?

If most of the clicks in the dataset reflect that generally, these ultimately lead to adding items to cart, perhaps this is some evidence that customers tend to know what they want when they land on the website. They aren't just querying the site or app out of general interest. 

### Creating a Ratings Method

Create the 1-2-3 scale by updating user_items dataset. Don't need to update the ActionId_Label.

#### Total Rating of Parent by Visitor 

Show the distribution of the new total rating score.

Compare it to other ratings transformations (i.e. lognormal, square root, coerce upper limit). 

#### Example of the Total Rating pre-processing pipeline.

walk through example records of the pre-processing steps for the Total Rating to illustrate the transformations.

### Setting up a User Ratings Matrix

## User Based Collaborative Filtering

## Item Based Collaborative Filtering