---
title: "EDA and Collaborative Filtering"
author: "Brandon Hoeft"
date: "October 22, 2017"
output: 
  html_document: 
    keep_md: yes # http://happygitwithr.com/rmd-test-drive.html#output-format
    toc: yes
    toc_depth: 3
---

```{r setup, include = FALSE}
# my global options defined for each code chunk.
knitr::opts_chunk$set(fig.width=7, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE, comment = '')
```

## Introduction

This Markdown file has most of my code, relevant plots, and explanations of decisions made so far in the data wrangling process.

I will update it with collaborative filtering model information too, as needed. 

## Getting the data into R

I used the following packages to do most of the initial data pre-processing and exploratory analysis. I'm working on an AWS AMI instance that's a linux machine with RStudio and git installed, using one of the instances publicly available and pre-configured from [Louis Aslett's website](http://www.louisaslett.com/RStudio_AMI/). I initially started with the free tier, *t2.micro* instance, but once I started the modeling phase, I scaled up to *m4.large*, which is also a general purpose computing instance but with better RAM, CPU, and better optimized memory management. AWS AMI instance costs can be calculated [here](https://calculator.s3.amazonaws.com/index.html). 

The datasets queried by Matt Hayden, I've put in my own S3 Bucket on Amazon Web Services (AWS). As such, the `aws.s3` [package](https://github.com/cloudyr/aws.s3) was used to access that bucket and bring the datasets into my environment. Access Keys are needed to make read/write calls to the bucket, but I've intentionally left that code out of here. 

``` {r echo = TRUE}

library(aws.s3) 
library(readr) # faster alternatives to base read. methods. 
library(lubridate) # date wrangling. 
library(dplyr) # data wrangling
library(tidyr) # data wrangling
library(ggplot2) # data viz. 
library(pryr) # mem_used() and object_size() functions to manage/understand memory usage.
```

To read the Part items data and the user webactivity data from s3 I passed read_table2() functions for reading tabular data into the `aws.s3` functions for reading objects from AWS. 

``` {r eval = FALSE, include = FALSE}
# specify personal account keys as environment variables so I can read my s3 object(s) from AWS. 

# DO NOT DO NOT DO NOT DO NOT DO NOT SAVE KEY in code or render in output!!!! Could compromise AWS account. 
#Sys.setenv("AWS_ACCESS_KEY_ID" = "",
#           "AWS_SECRET_ACCESS_KEY" = "")

```

``` {r echo = TRUE}

# items: part number, parent, catalogue, attributes/values.
items <- s3read_using(FUN = read_table2, 
                      object = "obfuscatedItems_10_17_17.txt", 
                      col_names = TRUE,
                      col_types = "cciciiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiiii",
                      bucket = "pred498team5")

# users: click data from the company website for a random day of user's selected, their activity for past 3 months and click summaries of how they interacted with parts. 
users <- s3read_using(FUN = read_table2, 
                      col_names = TRUE,
                      col_types = "ccici",
                      object = "obfuscatedWebActivity_10_16_17.txt", 
                      bucket = "pred498team5")
```

## Wrangling the Dataset 

The only information I'm particularly interested in from the **items** dataset is the Parent column, and left joining that to the matching PartNumber in the users data. Building a recommender system using the Parent family instead of the specific PartNumber will be discussed later. 

``` {r echo = TRUE}

# left join the parent family to the users PartNumber
user_items <- users %>%
  left_join(items, by = "PartNumber") %>%
  mutate(ActionDate = ymd(ActionDate), # parse into a date format.
         ActionId_label = factor(user_items$ActionId, # create labels
                                 labels = c("add to order", "select Part",
                                            "select Part detail",
                                            "print detail",
                                            "save CAD drawing detail",
                                            "print CAD drawing detail"))) %>% 
  select(-starts_with("Val"), -starts_with("Attr"), -starts_with("Catalog"))

# Check that left join worked and initial columns match identically. 
all.equal(users[1:3], user_items[1:3])

remove(users)
remove(items)
```

### Initial EDA

With a baseline dataset, I can start exploring some information about the users. I'll refer to them as visitors, since these are web click data about how visitors/customers of a manufacturing supplier are interacting with the company's website and mobile application. 

We'll start by asking some questions of the data. The web activity covered by these data begins on `r min(user_items$ActionDate)` and goes through `r max(user_items$ActionDate) `, covering a span of `r max(user_items$ActionDate) - min(user_items$ActionDate)` days. 

#### Summary of Unique Visitors, PartNumbers, Parents

Initial analysis shows that there are `r unique(user_items$VisitorId) %>% length()` unique VisitorID numbers in this dataset. 

There are `r unique(user_items$PartNumber) %>% length()` unique PartNumbers associated with the visitor web activity data.

Instead of looking at very specific PartNumbers, we can also roll the analysis up to the Parent level. The Parent level is a grouping of similar PartNumber products to a parent family. Within the website, it's a hyperlink associated with the PartNumber. Parent groups will vary. It is a way that individual merchandising managers at the company decide to categorize and split stuff up. There are `r unique(user_items$Parent) %>% length()` unique Parent family numbers (referred herein as Parents) associated with the visitor web activity data. 

Code for getting the aforementioned details is below. 

``` {r echo = TRUE, eval = FALSE }
# How many Different Users are in the dataset? 
unique(user_items$VisitorId) %>% length()
# How many distinct PartNumber's are in the dataset? 
unique(user_items$PartNumber) %>% length()
# How many distinct Parent Family's of PartNumbers are in the dataset? 
unique(user_items$Parent) %>% length()

```

#### How do total Action Counts of each visitor distribute? 

Action counts are how often a visitor did a particular action within their session. In a single session (the day of the visitor activity interaction with a PartNumber) if they selected a PartNumber in the carousel, this is reflected as a single row in the dataset, with an ActionCount of 1. If they selected the part 2 times in this session, the ActionCount value is equal to 2. 

``` {r echo = TRUE}
user_items %>%
  group_by(VisitorId) %>%
  summarize(activity_count = sum(ActionCount)) %>%
  with(summary(activity_count))

user_items %>%
  group_by(VisitorId) %>%
  summarize(activity_count = sum(ActionCount)) %>%
  ggplot(aes(activity_count, y = ..density..)) + 
  geom_histogram(binwidth = 100, colour = "black", fill = "darkgrey") +
  # median vertical line.
  geom_vline(aes(xintercept = median(activity_count)),
             color = "dodgerblue3", linetype = "dashed", size = 0.5) +
  geom_text(aes(0,.002, family = "courier", label = paste("median = ", median(activity_count))),
            nudge_x = 700, color = "dodgerblue3", size = 4.5) +
  labs(title = 'Distribution of Total interactions with PartNumbers',
       subtitle = 'For each Visitor')

```

After careful consideration, it was determined that total actions per PartNumber or per Parent should not influence the implicit rating derived for the recommender. This could be for a variety of reasons, such as selecting a part multiple times in the same session may perhaps be just as indicative of uncertainty as it could be of endorsing that part. 

#### How many unique PartNumbers are visitors interacting with? 

#### How many unique Parents are visitors interacting with? 

expect this to be less per user typically than the different PartNumbers 
they may interact with since Parent is a higher level grouping. 

#### How many unique visitors have interacted with each PartNumber? 

#### How many unique visitors have interacted with each Parent Family? 

Because Parent Family's are a broader generalization than PartNumbers, we'd 
expect to have more interactions within Parent Family than at the lower level
of PartNumber. 

#### How many PartNumbers fall into each Parent Family?

#### Do Parents with more PartNumbers also have more overall visitor clicks? 

#### What types of click Actions do the visitors take? 

### Creating a Ratings Method

Create the 1-2-3 scale by updating user_items dataset. Don't need to update the ActionId_Label.

#### Total Rating of Parent by Visitor 

#### Example of the Total Rating pre-processing pipeline.
